---
title: "GSS challange More Data"
author: "Kimberly Aldana"
date: "July 3 2023"
output: html_document
---







```{r include=FALSE,echo=FALSE}
source(file="fisherXY.R")
require(tidyverse)
require(tigerstats)
require(knitr)
require(foreign)
GSS <-read.dta(file="GSS2016.dta")
GSS <- droplevels(GSS)
mydata <- select(GSS,race,cappun)
mydata <- na.omit(mydata)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

Is attitude toward capitol punishment (cappun) independent of race (race)?
If not, then whose view differs from whom??

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

To analyze the relationship between "attitude toward capital punishment" (cappun) and "race," we can employ a combination of variable analysis, graphical displays, and inferential tools. A graphical display, such as a bar chart, can visually depict the distribution of attitudes toward capital punishment among different racial groups. Numerical methods, including cross-tabulation and percentage calculations, can provide a numerical summary of the relationship. Additionally, inferential tools such as the chi-square test and Fisher's exact test can be utilized to determine if there is a significant association between cappun and race. 



# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->


```{r}
barchartGC(~race + cappun,data=mydata)
barchartGC(~race + cappun,data=mydata, type="percent")

```
 
  Both plots aim to visualize the distribution of observations across different race and cappun (capital punishment) categories. The first plot shows the count of observations, while the second plot provides a relative comparison of proportions as percentages. Whites are the group that favors capital punishment in general, it indicates that a higher proportion or percentage of white individuals support the use of capital punishment compared to individuals from other racial groups. 



## Now with The Most data (fewer variables gives more data points)



### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~race + cappun, data=mydata)
rowPerc(table1)
colPerc(table1)
```

 These calculations help assess any patterns or disparities in attitudes towards capital punishment across different racial groups. The whites, black and other races all have an even amount of people in each race. White people are in favor of capital punishment with a percentage of 65.10%, also with having the lowest percentage of 34.90 of people who are opposed of capital punishment. Black people have the lowest percentage of 57.37 with the other races coming in 2nd with 44.80.



## Now with More data



## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

Something is going soon because the p value is low at .00000000000000022 so we need to run the fisher xy test. The p-value is still significant not as dramatic as the previous branch so we will still need to run a fisher xy test. Pearson's chi-squared test yielded a test statistic of 19.477 with 2 degrees of freedom, and a remarkably small p-value of 5.896e-05. This significant p-value provides strong evidence against the null hypothesis, suggesting that race and attitudes toward capital punishment are associated in the dataset.The two-sided alternative hypothesis implies that there is a discernible difference in attitudes toward capital punishment among the racial groups. 


```{r}
fisherXY(mydata$race,mydata$cappun)
```
Every race is different than with this particular data. Still having a very low p value. With the white race still holding the highest percentage in favor of capital punishment.This test evaluates whether there is a significant association between race and attitudes toward capital punishment based on the given dataset.The white and black race are different. There is no difference between white and other or black and other. The only p-value that is significant is between the white and black race, even if it had less data.


## More data version



# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
